---
title: "Getting Started with cdcmeasles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with cdcmeasles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The `cdcmeasles` package provides easy access to measles case data from the Centers for Disease Control and Prevention (CDC). This vignette will guide you through the basic usage of the package.

## Installation

You can install the development version from GitHub:

```r
# install.packages("devtools")
devtools::install_github("dwchal/cdcmeasles")
```

## Loading the Package

```{r load}
library(cdcmeasles)
```

## Checking Data Availability

Before downloading data, you can check if the CDC data endpoints are accessible:

```{r availability, eval=FALSE}
# Check availability (returns named logical vector)
is_data_available()

# With detailed output
is_data_available(verbose = TRUE)
```

## Understanding the Data Sources

The package accesses two official CDC data sources:

1. **Weekly data**: Recent case counts aggregated by week
2. **Yearly data**: Historical case counts with outbreak information

You can view the metadata and URLs:

```{r metadata}
# Get package metadata
meta <- get_measles_metadata()
print(meta$data_urls)
print(meta$weekly_columns)
print(meta$yearly_columns)
```

## Downloading Data

### Weekly Data

Weekly data contains recent measles case counts:

```{r weekly, eval=FALSE}
# Download weekly data
weekly <- get_measles_data("weekly")
head(weekly)
```

Expected columns:
- `week_start`: Start date of the reporting week
- `week_end`: End date of the reporting week
- `cases`: Number of reported cases

### Yearly Data

Yearly data contains historical case counts with additional context:

```{r yearly, eval=FALSE}
# Download yearly data
yearly <- get_measles_data("yearly")
head(yearly)
```

Expected columns:
- `year`: The year
- `cases`: Total cases for the year
- `states_with_cases`: Number of states reporting cases
- `outbreaks`: Description of major outbreaks

## Quick Summary

Get a quick summary of the most recent data:

```{r summary, eval=FALSE}
summary <- get_current_year_summary()
print(summary)
```

## Visualizations

The package includes several visualization functions that use ggplot2.

### Time Series Plot

```{r timeseries, eval=FALSE}
# Weekly time series
weekly <- get_measles_data("weekly")
plot_measles_time_series(weekly)

# Yearly time series with customization
yearly <- get_measles_data("yearly")
plot_measles_time_series(
  yearly,
  date_col = "year",
  title = "Historical Measles Cases in the US",
  line_color = "steelblue",
  show_points = TRUE
)
```

### Bar Chart

```{r barchart, eval=FALSE}
yearly <- get_measles_data("yearly")
plot_measles_bars(
  yearly,
  title = "Measles Cases by Year",
  highlight_recent = TRUE
)
```

### State Map

For state-level data (if available), you can create choropleth maps:

```{r map, eval=FALSE}
# Example with sample data
state_data <- data.frame(
  state = c("california", "texas", "new york", "florida"),
  cases = c(50, 30, 45, 20)
)

plot_measles_state_map(state_data)
```

Note: The `maps` package is required for map visualizations.

## Data Cleaning

The package provides a utility function for data cleaning:

```{r cleaning, eval=FALSE}
# Clean data and remove rows with NA cases
weekly <- get_measles_data("weekly")
clean_weekly <- clean_measles_data(weekly, remove_na = TRUE)
```

## Error Handling

The package handles network errors gracefully:

```{r errors, eval=FALSE}
# With verbose output to see what's happening
data <- get_measles_data("weekly", verbose = TRUE)

# Returns NULL if download fails
if (is.null(data)) {
  message("Data not available")
}
```

## Working Offline

You can save downloaded data for offline use:

```{r offline, eval=FALSE}
# Download and save
weekly <- get_measles_data("weekly")
write.csv(weekly, "measles_weekly.csv", row.names = FALSE)

# Later, read back
weekly <- read.csv("measles_weekly.csv")
weekly$week_start <- as.Date(weekly$week_start)
weekly$week_end <- as.Date(weekly$week_end)
```

## Notes on Data Access

- The CDC may update their data structure or URLs periodically
- Use `get_cdc_urls()` to check the current endpoints
- Network connectivity is required for downloading data
- Consider caching data locally for repeated analyses

## Further Resources

- [CDC Measles Data Page](https://www.cdc.gov/measles/data-research/index.html)
- [Package GitHub Repository](https://github.com/dwchal/cdcmeasles)
