---
title: "Introduction to cdcmeasles"
author: "Your Name"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to cdcmeasles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The `cdcmeasles` package provides an easy way to access measles data from the Centers for Disease Control and Prevention (CDC). This vignette demonstrates the basic functionality of the package.

## Installation

You can install the development version of `cdcmeasles` from GitHub:

```{r installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("yourusername/cdcmeasles")
```

## Basic Usage

First, load the package:

```{r setup}
library(cdcmeasles)
```

### Checking Data Availability

Before downloading data, we can check if the CDC data sources are available:

```{r check-availability}
is_data_available()
```

### Getting Metadata

We can retrieve metadata about the dataset:

```{r metadata}
get_measles_metadata()
```

### Downloading Data

The main function of the package is `get_measles_data()`, which downloads measles data from the CDC:

```{r download-data, eval=FALSE}
# This chunk is set to eval=FALSE to prevent actual data download during vignette building
measles_data <- get_measles_data()
head(measles_data)
```

You can also save the data directly to a CSV file:

```{r download-save, eval=FALSE}
measles_data <- get_measles_data(save_file = TRUE, file_name = "measles_data.csv")
```

## Data Visualization

The package includes functions to visualize the data:

### Time Series Plot

```{r time-series, eval=FALSE}
# Assuming the data has appropriate date and cases columns
plot_measles_time_series(measles_data)
```

### State Map

If the data includes state-level information:

```{r state-map, eval=FALSE}
# Assuming the data has appropriate state and cases columns
plot_measles_state_map(measles_data)
```

## Working with the Data

Once you have downloaded the data, you can analyze it using standard R functions and packages:

```{r data-analysis, eval=FALSE}
# Summary statistics
summary(measles_data)

# If using dplyr for data manipulation
if (requireNamespace("dplyr", quietly = TRUE)) {
  library(dplyr)
  
  # Example: Find the states with the highest number of cases
  # (Adjust column names based on actual data structure)
  measles_data %>%
    group_by(state) %>%
    summarize(total_cases = sum(cases, na.rm = TRUE)) %>%
    arrange(desc(total_cases)) %>%
    head(10)
}
```

## Conclusion

The `cdcmeasles` package provides a convenient interface to CDC measles data, making it easier for researchers, public health professionals, and data scientists to access and analyze this important public health information.

Remember to check the CDC website for information about data updates and changes to the data structure:
https://www.cdc.gov/measles/data-research/index.html 